---
title: "How to Read, Write, and Google for Success in Coding"
author: |
        | Christopher Skovron
        | Northwestern University 
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  revealjs::revealjs_presentation: 
    theme: simple
    highlight: haddock
    center: false
    transition: none
    #css: reveal.css
    self_contained: false
---
          
          
## You probably hear a lot about "clean" code 

## Most of that is bullshit and pandering
- unless you are working in industry in a setting where things need to highly scale 

## In five years, you will look back on your past self's code and be horrified 

## Reading and writing documents related to coding
- You will be reading new kinds of documentation, vignettes, examples, error messages
- These will seem confusing at first
- There are strategies, bu tnothing can beat practice 

## Your goals in writing code are more modest 
- Does this do what I need it to do in a reasonably computationally efficient manner? 
- Will my future self understand it? 
- Will my collaborators and/or replicators understand it? 
- Have I written in such a way that will prevent me from making mistakes? 
We will address each of these goals in order. 



# How to comment code/being kind to your future self 






# Some resources to orient you to best practicies in writing code

- Read points 1-8 of [this guide](https://www.r-bloggers.com/r-best-practices-r-you-writing-the-r-way/) (specific to R but applies to other languages)
- [Hadley Wickham's R style guide](http://adv-r.had.co.nz/Style.html)
- [Another good resource on R best practices.](https://swcarpentry.github.io/r-novice-inflammation/06-best-practices-R/)








## Pick good names
- You might be tempted to use abbreviations or non-words to save yourself from typing the same characters over and over
- With modern IDEs, tab completion means this is not a big issue
- Choose names that are legible, clear, and precise 
- Name functions what they do, name variables what they are 


## Good and bad names

```{r, eval=FALSE}
## a few bad variable names from my past work
incspendtran
gender # coded as one or zero


## better names
increase.spending.transportation
female # coded 1 if female


```

## Similar variables should have similar names

- For example, my candidates' perceptions of things alwas begin with `perc.`
- Next, geography, are they perceiving district, `perc.district`, or state, `perc.state`
- Sometimes they are perceiving opinion within a party: `perc.district.dem`
- Then the issue: `perc.district.dem.marriage` means their district-level perception of support for same-sex marriage. `perc.district.guns.banassault` is district-level support for banning assault weapons. There's also `guns.background.check`....




## Principles of good commenting 

- Don't comment things that are painfully obvious
- But most things won't be painfully obvious 
- Comment to introduce each new section of code or each major task 
- If you write a new function 






# How to find help

## Expect to constantly be searching for help 

- There's way too much to remember, we offload that knowledge to the internet and elsewhere
- 



## Do due diligence before googling aimlessly

- You may be able to google a solution, but that is sort of the being-given-a-fish way. You want to learn to fish. 


# How to read documentation





## Build up a library of cheats 

- Bookmark useful help pages
- Make yourself a notebook of examples of code. Have a dropbox folder where you save snippets that do tasks you frequently do. 
- I have things like group-level proportions that I frequently do for summary statistics 
- Pillage old code for new projects (But be careful! You or your software have probably improved, so writing it anew could be a good learning experience and/or improve efficiency)



## Someone else has had your problem before 
- You will almost never need to post your own question to a mailing list or to StackOverflow. 
- Someone has had a similar problem before, you just need to know how to find it
- People might have similar but slightly different problems that you can 


## Use built-in help

```{r}
?lm
?read.csv
?dplyr::rename
```

## Built-in help will
- Tell you how functions work, to varying degrees of helpfulness 
- Often provide some examples of how functions work
- Often still leave you clueless and needing to google around some more 

## Figuring out your google query




## Asking for help
- You will want a "minimal workable example"



